---
title: "Untitled"
output: html_document
---

#Individual Scoring Functions
#score_rf
```{r, echo=FALSE}
RF_Score<-function(data, response, explanatory){
  library(pROC)
  library(hydroGOF)
  library(randomForest)
  
  set.seed(0)
  
  results<-NULL
  
  rf<-randomForest(as.factor(data[,response])~., data=data[explanatory])
  score<-auc(data[,response], as.numeric(predict(rf)))
  loop_results<-cbind('Model'='Random Forest', 'Response'=response, 'Explanatory'=list(explanatory), 'Score_Type'='AUC', 'Score'=score)
    
  results<-data.frame(rbind(loop_results, results))
  
  return(results)
}
```
#score_dt
```{r, echo=FALSE}
DT_Score<-function(data, response, explanatory){
  library(pROC)
  library(hydroGOF)
  library(rpart)
  
  set.seed(0)
  
  results<-NULL
  
  dt<-rpart(as.factor(data[,response])~., data=data[explanatory])
  score<-auc(data[,response], as.numeric(predict(dt)[,1]))
  loop_results<-cbind('Model'='Decision Tree', 'Response'=response, 'Explanatory'=list(explanatory), 'Score_Type'='AUC', 'Score'=score)

  results<-data.frame(rbind(loop_results, results))
  
  return(results)
}
```
#score_nb
```{r, echo=FALSE}
NB_Score<-function(data, response, explanatory){
  library(pROC)
  library(hydroGOF)
  library(klaR)
  
  set.seed(0)
  
  results<-NULL
  
  nb<-NaiveBayes(as.factor(data[,response])~., data=data[explanatory])
  score<-auc(data[,response], as.numeric(unlist(predict(nb)[1])))
  loop_results<-cbind('Model'='Naive Bayes', 'Response'=response, 'Explanatory'=list(explanatory), 'Score_Type'='AUC', 'Score'=score)
    
  results<-data.frame(rbind(loop_results, results))
  
  return(results)
}
```
#score_sv
```{r, echo=FALSE}
SV_Score<-function(data, response, explanatory){
  library(pROC)
  library(hydroGOF)
  library(e1071)
  
  set.seed(0)
  
  results<-NULL
  
  sv<-svm(as.factor(data[,response])~., data=data[explanatory])
  score<-auc(data[,response], as.numeric(predict(sv)))
  loop_results<-cbind('Model'='Support Vector Machine', 'Response'=response, 'Explanatory'=list(explanatory), 'Score_Type'='AUC', 'Score'=score)
  
  results<-data.frame(rbind(loop_results, results))
  
  return(results)
}
```
#Combined Scoring Functions
```{r, echo=FALSE}
Scoring_Results<-function(data, response, explanatory){
  total_results<-NULL
  top_scores<-NULL
  
  total_results<-data.frame(rbind(RF_Score(data, response, explanatory),  DT_Score(data, response, explanatory),  NB_Score(data, response, explanatory),  SV_Score(data, response, explanatory)))
  
  total_results$Score<-as.numeric(as.character(total_results$Score))
  type<-unique(total_results$Score_Type)

  total_results[order(-total_results$Score),]

  return(total_results)
}
```
#Individual Prediction Functions
#predict_rf
```{r, echo=FALSE}
RF_Predict<-function(data, response, explanatory){
  rf<-randomForest(as.factor(data[,response])~., data=data[explanatory])
  data[, paste(response,"_Pred", sep="")]<-predict(rf)
}
```
#predict_dt
```{r, echo=FALSE}
DT_Predict<-function(data, response, explanatory){
  dt<-rpart(as.factor(data[,response])~., data=data[explanatory])  
  data[, paste(response,"_Pred", sep="")]<-predict(dt)
}
```
#predict_nb
```{r, echo=FALSE}
NB_Predict<-function(data, response, explanatory){
  nb<-NaiveBayes(as.factor(data[,response])~., data=data[explanatory])
  data[, paste(response,"_Pred", sep="")]<-predict(nb)  
}
```
#predict_sv
```{r, echo=FALSE}
SV_Predict<-function(data, response, explanatory){
  sv<-svm(as.factor(data[,response])~., data=data[explanatory])
  data[, paste(response,"_Pred", sep="")]<-predict(sv)  
}
```
#Combined Prediction Functions
```{r, echo=FALSE}
Prediction_Results<-function(model, data, response, explanatory){
  if (model=='Random Forest'){
    results<-RF_Predict(data, response, explanatory)
  } else if (model=='Decision Tree'){
    results<-DT_Predict(data, response, explanatory)
  } else if (model=='Naive Bayes'){
    results<-NB_Predict(data, response, explanatory)
  } else if (model=='Support Vector Machine'){
    results<-SV_Predict(data, response, explanatory)
  }
  return (results)
}
```
#Combined Scoring and Prediction Functions
```{r, echo=FALSE}
Combined_Results<-function(data, response, explanatory){
  score_results<-Scoring_Results(data, response, explanatory)

  #print(score_results)
  
  type<-unique(score_results$Score_Type)

  top_results<-droplevels(score_results[which.max(score_results$Score), ])
  
  print(top_results)
  #return(top_results)
  
  prediction_results<-Prediction_Results(top_results$Model, data, response, explanatory)
  
  return(prediction_results)
}
```

#Load Raw Data and Libraries
```{r,echo=FALSE}
options(warn=-1)

Attacks <- read.csv("C:/Users/bethany.earnest/Desktop/Harvard/Spring2016/Shark_Attack_Data_4-7-2016.csv")
Activities<-read.csv("C:/Users/bethany.earnest/Desktop/Harvard/Spring2016/Project/Sam_Activities_4-30-2016.csv")
Species<-read.csv("C:/Users/bethany.earnest/Desktop/Harvard/Spring2016/Project/Sam_Species_4-30-2016.csv")
Temp<-read.csv("C:/Users/bethany.earnest/Desktop/Harvard/Spring2016/Project/Ruth_Temp_5-1-2016.csv")

Activities$Case.Number<-Activities$CaseNumber
Activities$CaseNumber<-NULL

Species$Case.Number<-Species$CaseNumber
Species$CaseNumber<-NULL

Temp$Case.Number<-Temp$CaseNumber
Temp$CaseNumber<-NULL

library(dplyr)

Attacks<-left_join(Attacks, Activities, by="Case.Number")
rm(Activities)
Attacks<-left_join(Attacks, Species, by="Case.Number")
rm(Species)
Attacks<-left_join(Attacks, Temp, by="Case.Number")
rm(Temp)

Attacks<-droplevels(subset(Attacks, 
                           Country=="USA" &
                            Year>1980))
```

#Wrangle Data
#Fatal [BETH]
```{r,echo=FALSE}
Attacks$Fatal<-Attacks$Fatal..Y.N.
Attacks$Fatal..Y.N.<-NULL

table(Attacks$Fatal)

Attacks<-droplevels(subset(Attacks, 
                              !Fatal %in% c("F", "UNKNOWN","") | 
                                is.na(Fatal)))

table(Attacks$Fatal)

for(i in 1:nrow(Attacks)){
  for(n in unique(Attacks$Fatal)){
    Attacks[i,]$Fatal<-ifelse(Attacks[i,]$Fatal!="Y", "N", "Y") 
  }
}

table(Attacks$Fatal)

Attacks<-droplevels(Attacks)

table(Attacks$Fatal)

Attacks$Fatal_Binary<-as.numeric(Attacks$Fatal)
Attacks$Fatal_Binary<-Attacks$Fatal_Binary-1
```
#Age [BETH]
```{r,echo=FALSE}
#table(Attacks$Age)

Attacks$Age<-as.numeric(Attacks$Age)

hist(Attacks$Age)

Attacks<-droplevels(subset(Attacks, Age>1))

hist(Attacks$Age)
```
#Month [BETH]
```{r,echo=FALSE}
Attacks$Clean_Date<-as.Date(Attacks$Date, format="%d-%b-%Y")

Attacks$Month<-format(Attacks$Clean_Date, format="%B")

table(is.na(Attacks$Month))

Attacks<-droplevels(subset(Attacks, !is.na(Month)))

table(is.na(Attacks$Month))

Attacks$Clean_Date<-NULL
```
#Gender [BETH]
```{r,echo=FALSE}
table(Attacks$Sex)

Attacks<-droplevels(subset(Attacks, 
                              !Sex %in% c("lli") | 
                                is.na(Sex)))

table(Attacks$Sex)

for(i in 1:nrow(Attacks)){
  for(n in unique(Attacks$Sex)){
    Attacks[i,]$Sex<-ifelse(Attacks[i,]$Sex!="F", "M", "F") 
  }
}

table(Attacks$Sex)

Attacks<-droplevels(Attacks)

table(Attacks$Sex)
```

#Species [SAM]
#Activity [SAM]
[Activity (Sam)](http://something_blah.html)
[Activity (Sam)](something_blah.html)
#Temp [RUTH]

#Address Missing Values
```{r,echo=FALSE}
for(i in names(Attacks)){
  print(names(Attacks[i]))
  print(table(is.na(Attacks[,i])))
  
  #if(names(Attacks[i])=="activity_cat"){
  #  library(randomForest)
  #  Attacks[,i]<-na.roughfix(Attacks[,i])
  #}
  if(class(Attacks[,i])=="character"){
    Attacks[,i]<-as.factor(Attacks[,i])
  }
}
```

#Binary Classification
#Define Variables for Model
```{r,echo=FALSE}
library(randomForest)
Attacks<-na.roughfix(Attacks)

explanatory_vars<- c("Month","Sex","Age","Type")
Combined_Results(Attacks, "Fatal_Binary", explanatory_vars)

explanatory_vars<- c("Month","Area","Sex","Age","Type")
Combined_Results(Attacks, "Fatal_Binary", explanatory_vars)

explanatory_vars<- c("Month","Area","Sex","Age","Type", "activity_cat")
Combined_Results(Attacks, "Fatal_Binary", explanatory_vars)

explanatory_vars<- c("Month","Area","Sex","Age","Type", "activity_cat", "species_cat")
Combined_Results(Attacks, "Fatal_Binary", explanatory_vars)

explanatory_vars<- c("Month","Area","Sex","Age","Type", "activity_cat", "species_cat", "temp")
Combined_Results(Attacks, "Fatal_Binary", explanatory_vars)
```
#Results
```{r,echo=FALSE}
Monthly_Results<-Attacks %>% group_by(Month) %>% summarize(Probability=mean(Fatal_Binary)) %>% arrange(desc(Probability))
Gender_Results<-Attacks %>% group_by(Month, Sex) %>% summarize(Probability=mean(Fatal_Binary)) %>% arrange(desc(Probability))

Female_Results<-Attacks %>% filter(Sex=='F') %>% group_by(Month, activity_cat) %>% summarize(Probability=mean(Fatal_Binary)) %>% arrange(desc(Probability))

Male_Results<-Attacks %>% filter(Sex=='M') %>% group_by(Month, activity_cat) %>% summarize(Probability=mean(Fatal_Binary)) %>% arrange(desc(Probability))

Activity_Results<-Attacks %>% group_by(activity_cat) %>% summarize(Probability=mean(Fatal_Binary)) %>% arrange(desc(Probability))

setwd("C:/Users/bethany.earnest/Desktop/Harvard/Spring2016/Project/")
```