---
title: "SharkAttack"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following data sets are provided to analyze Shark Attack.

```{r}
# http://www.huffingtonpost.com/norm-schriever/coming-down-from-shark-we_b_3740495.html
suppressMessages(library(readr))
suppressMessages(library(tidyr))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
# attack location, how many people in the area, water realted env variables, date, time
SAdata <- read_csv("Shark_Attack_Data_4-7-2016.csv",col_names=TRUE,n_max=5897)
SharkSpecies <- tbl_df(read_csv("Shark_Species_Data_4-7-2016.csv"))
Sharks <- read_csv("Sharks_4-7-2016.csv")
```

## Analyze area, location

Note that "Unprovoked attacks" are defined as incidents where an attack on a live human occurs in the shark’s natural habitat with no human provocation of the shark.

```{r}
SA_tab <- SAdata[1:19] 
SA_tab %>% select(`Case Number`,Date,Year,Location,Area,Country) %>% 
  group_by(Location,Area,Country) %>% summarize(n=n()) %>% 
  filter(is.na(Area)) %>% View()
SA_tab %>% group_by(Location) %>% select(`Case Number`,Date,Year,Area) %>%
  summarize(n=n()) %>% print(n=3838)
```

# include plots

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
