---
title: "Wrangling Shark Attack Data: Species & Activity"
author: "Fisher"
date: "April 13, 2016"
output: html_document
---

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(readr)
library(dplyr)
library(stringr)
library(stringi)
library(knitr)
library(wordcloud)
library(gridExtra)
```

```{r}
# Load the data 
dat <- read_csv("Shark_Attack_Data_4-7-2016.csv",col_names=TRUE,n_max=5897) #5897=last good row
dat[24] <- NULL #remove null rows and repeated Case Number rows
dat[23] <- NULL
dat[21] <- NULL
dat[20] <- NULL
```

```{r,eval=FALSE}
# Working with Windows to Unicode special characters...

#dat %>% filter(grepl("(Bay, near Charleston)",Location)) %>% select(Location) %>% #View()
#  mutate(Location=iconv(Location,from="CP1252",to="UTF-8")) %>% View()
#  #filter(grepl("([’])",Location)) %>% View()
#dat %>% filter(grepl("(Bay, near Charleston)",Location)) %>% select(Location) %>% 
#  mutate(Location=stri_escape_unicode(iconv(Location,from="CP1252",to="UTF-8"))) %>% 
#  #mutate(Location=gsub("’","",Location,perl=TRUE)) %>% View()
#  mutate(Location=gsub("(\\\\u2019)","'",Location,perl=TRUE)) %>% View()
```

##Wrangle the Species column
The Species variable provides textual descriptions of the species of shark reported in the shark attack. The descriptions are free form. In addition to possibly including the **type of shark**, many entries include its **size** (length and/or weight range), **gender**, and other miscellaneous information. Therefore, the Species variable needs to be wrangled, or normalized, so that the shark types can be parsed and extracted. As the additional information contained in a Species entry may be useful for downstream processing and analysis, it is also wrangled and used to create additional variables, e.g., length, weight, and gender.

Note, only `r dat %>% filter(!is.na(Species)) %>% nrow()` of `r nrow(dat)` reported incidents contain information about the shark species.

####Normalize the Species descriptions

General normalization...
```{r}
## Normalize all Species column entries

dat2 <- dat %>% filter(!is.na(Species)) %>% 
  mutate(Species=iconv(Species,from="CP1252",to="UTF-8"),
         Species=stri_trans_tolower(Species)) %>% 
  select(`Case Number`,Species) # Lowercase all Species text

dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub("([0-9]{1,2}\\/[0-9]{1,2}\\/[0-9]{4})"," ",Species))
# Remove annoying unicode characters
dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub("(\u393C|\u3E33|\u613C|\u3E30|\u3E34|\u623C|\u3E64|\uFFFD)","",Species))

## Remove all double quotes at the beginning and end of each Species entry
dat2 <- dat2 %>% filter(!is.na(Species)) %>% 
  mutate(Species=gsub(pattern="(^|[\\s])[\"]"," ",Species), #from beginning  “
         #Species=gsub(pattern="([^0-9])([\"])([\\S]|$)","\\1",Species)) #ending (unless preceded by a digit)
         Species=gsub(pattern="([\"])([a-z -]+)([\"])","\\2",Species),
         Species=gsub(pattern="([a-z -]+)([\"])","\\1",Species))

dat2 <- dat2 %>% filter(!is.na(Species)) %>% #Remove unnecessary punctuation
  mutate(Species=gsub(pattern="([a-zA-Z])[\\.]","\\1 ",Species), #removes word-final periods
         Species=gsub(pattern="\\?+"," ",Species), #removes question marks
         Species=gsub(pattern="[,;:]"," ",Species), #removes commas & semicolons
         Species=gsub(pattern="\\[|\\]"," ",Species), #removes square brackets
         Species=gsub(pattern="\\(|\\)"," ",Species)) #, #remove parentheses
         #Species=gsub(pattern="([e])[\"][\\s]","\\1 ",Species)) #strange double quote won't go away.


dat2 <- dat2 %>% filter(!is.na(Species)) %>% #normalize shark length descriptions
  mutate(Species=gsub(pattern="pound","lb",Species), #pound to lb
    Species=gsub(pattern="(\\d)(\\s|-)(metre|m|meter|cm|ft|feet|foot|kg|lb|lbs)(\\s|$)","\\1\\3 ",Species),
         Species=gsub(pattern="(\\d)(?-)(ft|feet|foot)(\\s|$)","\\1' ",Species),
         Species=gsub(pattern="(\\d)(?-)(meter|metre)(\\s|$)","\\1m ",Species)) 

dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub(pattern="(\\w)(\\s|-)(tipped)(\\s|$|\")","\\1tip ",Species), #normalizes tipped to tip
         Species=gsub(pattern="(sharks|shar k|shar ;)","shark",Species), # normalize 'shark'
         Species=gsub(pattern="( shar)\\s+([^k])"," shark \\2",Species), # normalize 'shark'
         
         Species=gsub(pattern="(\\d)(')(shark)","\\1\\2 \\3",Species), #5'shark to 5' shark
         Species=gsub(pattern=" (\")(\\w)"," \\2",Species), #turns "word into word
         Species=gsub(pattern="(fragment|wobbegong|day|pointer)(s)","\\1",Species), #plural to singular
         Species=gsub(pattern=" & "," and ",Species), # & to and
         Species=gsub(pattern="zambezi","zambesi",Species), #zambesi
         Species=gsub(pattern="grey","gray",Species), #gray
         Species=gsub(pattern="(seven gill|seven-gill|7-gill)","sevengill",Species), # sevengill
         Species=gsub(pattern="(sandtiger|sand-tiger)","sand tiger",Species), # sand tiger
         Species=gsub(pattern="([\\s]nose|-nosed|[\\s]nosed)","nose",Species)) # broadnose, ..nose

dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub(pattern="catshark","cat shark",Species))

dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub("(not confirmed|not cofirmed)","unconfirmed",Species))

dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub("(not authenticated)","unauthenticated",Species))


#Remove stop words from entries
#dat2 <- dat2 %>% filter(!is.na(Species)) %>%
#  mutate(Species=gsub("(^|\\s)(another|been|a|as|an|by|of|or|on|that|his|but|it|its|their|than|from|have|out|were|be|is|was|later|on|with|and|the|')($|\\s)"," ",Species))
dat2 <- dat2 %>% filter(!is.na(Species)) %>%
  mutate(Species=gsub("(^|\\s)(another|been|a|as|an|by|of|or|on|that|his|but|it|its|their|than|from|have|out|were|be|is|was|later|on|with|and|the)($|\\s)"," ",Species))

dat2 <- dat2 %>% filter(!is.na(Species)) %>% 
  mutate(Species=gsub(pattern="\\s+"," ",Species)) #Reduce multiple spaces to 1 space

#dat2 %>% filter(!is.na(Species)) %>% select(Species) %>% View()
#dat2$Species[675]
```

The following word clouds illustrate the textual information contained in the Species column. 
```{r,echo=FALSE}
species_texts <- dat2 %>% filter(!is.na(Species)) %>% select(Species)
species_text <- paste(unlist(species_texts),sep=" ",collapse = " ")
#species_text <- gsub(" [\\.] "," ",species_text)
#species_text <- gsub("[\\p{Cc}]","",species_text,perl=TRUE) # remove cntrl characters

species_tokens <- unlist(str_split(species_text," "))
species_token_counts <- as.data.frame(table(species_tokens)) %>% filter(nchar(as.character(species_tokens))>2)
#species_token_counts %>% View()

# Show the words relative to frequency
wordcloud(species_token_counts$species_tokens,
          species_token_counts$Freq,min.freq = 4,#c(4,1),
          random.order=FALSE, 
          colors=brewer.pal(8, "Dark2"))

cat("Leaving out the most frequent word 'shark'...",sep="")
tmp <- species_token_counts %>% filter(species_tokens!="shark")
wordcloud(tmp$species_tokens,tmp$Freq,
            min.freq = 4,#c(4,1),
            random.order=FALSE, 
            colors=brewer.pal(8, "Dark2"))

#gridExtra::grid.arrange(wc1,wc2,nrow=1)

# write the species tokens out to csv file.
#write.csv(species_token_counts,"species_column_tokens.csv",row.names = FALSE,quote=FALSE)

# From species_tokens, identify the morphological characteristics 
# of size (length, weight) and gender descriptions


```



```{r,eval=FALSE,echo=FALSE}
shark_texts <- species_texts %>% 
  mutate(shark.type=ifelse(grepl("shark",Species,fixed=TRUE),
                        str_extract_all(Species,"([^\\d\\s-']+[a-z]{2,9}[\\s]+){1,2}(shark)"),
                           Species),
         shark.type=ifelse(shark.type=="character(0)",NA,shark.type))

shark_texts <- shark_texts %>% #filter(is.na(shark.type)) %>% 
  mutate(shark.type=ifelse(is.na(shark.type),paste0(str_extract_all(Species,"(shovelnose|guitarfish|white|tiger|bull|shortfin|mako|lemon|oceanic|whitetip|blue|bluenose|blue pointer|galapagos|galagapos|caribbean|reef|dusky|blacktip|silky|gray|reef|great|hammerhead|blacktip|reef|broadnose|sevengill|bluntnose|sixgill|tawney|tawny|nurse|sand|tiger|spotted|catshark|cat|wobbegong|basking|spinner|bronze whaler|bronze|whaler|carchar[^\\s]+)"),sep=" "),shark.type),
         shark.type=ifelse(shark.type=="character(0)",NA,shark.type))

shark_texts <- shark_texts %>% 
  mutate(shark.type=ifelse(grepl("c\\(",shark.type),gsub("(c\\(|[,\"]|\\))","",shark.type),shark.type))

shark_texts <- shark_texts %>% 
  mutate(shark.type=ifelse(grepl("character\\(0\\)",shark.type),NA,shark.type))         
tmp <- shark_texts %>% filter(!is.na(shark.type)) %>% select(shark.type) 

tmp <- sort(unlist(tmp)) #%>% sort(shark.type)

removeStuff <- function(s){
  s <- gsub("(^|[\\s]+)(to|\\/|c|in|authenticated|recovered|fragment|species|attack|reported|not|female|remains|likely|the|after|aka|recovered|caused|cold|description|diameter|dr|dried|probable|incident|believed|identified|involve|possibly|little|small|juvenile|large|[a-z]|[\\.\\d\'\"]+|[\\d][\\.]?[\\dmc]{1,2})([\\s]|$)"," ",s, perl = TRUE)
  s <- gsub("[\\s]+"," ",s, perl = TRUE)
  (str_trim(s))
}

#shark_names <- sapply(tmp$shark.type,FUN=removeStuff) %>% table() %>% as.data.frame()
shark_names <- sapply(tmp,FUN=removeStuff) %>% table() %>% as.data.frame()
colnames(shark_names) <- c("name","freq")
shark_names <- shark_names[-1,] %>% filter(name!="shark")
#shark_names %>% View()


#write.csv(shark_names,file = "shark_names_1.csv",row.names = FALSE)

#shark_names %>% arrange(desc(Freq)) %>% head(20)
#shark_names %>% arrange(desc(Freq)) %>% tail(20)

wordcloud(shark_names$name,shark_names$freq,
            min.freq = 2,#c(4,1),
            #random.order=FALSE, 
            colors=brewer.pal(8, "Dark2"))

shark_names_text <- str_trim(paste(shark_names$name,collapse=" "))
shark_name_words <- as.data.frame(table(w=str_split(shark_names_text," ")))
shark_name_words <- shark_name_words %>% filter(w!="shark")
wordcloud(shark_name_words$w,shark_name_words$Freq,
            min.freq = 3,c(2.5,1),
            #random.order=FALSE, 
            colors=brewer.pal(6, "Dark2"))
```

As illustrated, further wrangling is needed. The word clouds show a variety of information that is not in itself categorically coherent. 

##Extracting from the Species Column
After exploring, normalizing, analyzing, and visualizing the species entry descriptions, the next step is to extract the three categories. Let's extract the Species descriptions into three distinct categories: (1) Shark Type, describing the type of shark; (2) Size, describing the size of the shark; and (3) Gender. 

####Shark Type (aka Species)
Here, we'll add a new column (variable) containing textual descriptions of the species of shark in an attack. This will be cross-correlated with the Shark Species data set.

```{r}
# sharky words..

#shark_type_pattern <- "(zygena|zygaena|albimarginatus|bamboo|blackfin|blacktip|blue|blue fin|bluefin|bluenose|nose|bonito|bonnethed|bronze|brown|brown-colored|bull|canicula|carcharhinid|carcharhinus|carchar[^\\s]+|caribbean|cat |catshark|cocktail|copper|cow|dog|dusky|francisci|galapagensis|galapagos|galap[^\\s]+|galeolamna|gangeticus|gray|ground|guitarfish|gummy|hammerhead|heterodontus|hooked shot|horn|lemon|leopard|leucas|long thin brown-colored|longfin|loxodon|maculipinnis|mako|nurse|obscurus|oceanic|perezi|porbeagle|raggedtooth|reef|rhizoprionodon|salmon|sand|sandbar|scalloped|scyliorhinus|sevengill|shark|shortfin|shovelnose|silky|silvertip|sixgill|smooth|soupfin|spinner|spotted|spurdog|tawney|tawny|thresher|tintorera|tintorero|whaler|white|whitetip|whtietip|wobbegong|zambesi)"

shark_type_terms <- c("zygena", "zygaena", "albimarginatus", "bamboo", "blackfin", "blacktip", "blue", "blue fin", "bluefin", "bluenose", "nose", "bonito", "bonnethed", "bronze", "brown", "brown-colored", "bull", "canicula", "carcharinid", "carcharhinid", "carcharhinus", "caribbean", "cat ", "catshark", "cocktail", "copper", "cow", "dog  ", "dusky", "francisci", "galapagensis", "galapagos", "galap[^\\s]+|galeolamna", "gangeticus", "gray", "ground", "guitarfish", "gummy", "hammerhead", "heterodontus", "hooked shot", "horn", "lemon", "leopard", "leucas", "long thin brown-colored", "longfin", "loxodon", "maculipinnis", "mako", "nurse", "obscurus", "oceanic", "perezi", "pointer","porbeagle", "raggedtooth", "reef", "rhizoprionodon", "salmon", "sand", "sandbar", "scalloped", "scyliorhinus", "sevengill", "shark", "shortfin", "shovelnose", "silky", "silvertip", "sixgill", "smooth", "soupfin", "spinner", "spotted", "spurdog", "tawney", "tawny", "thresher", "tiger","tintorera", "tintorero", "whaler", "white", "whitetip", "whtietip", "wobbegong", "zambesi", "no shark")

# Remove non-descriptive words..
removeNonTypeWords <- function(s){
  s <- gsub("(^| )(involvement|unconfirmed|description|allegedly|identified|recovered|according|inflicted|reported|diameter|fragment|probable|incident|benjamin|believed|juvenile|possibly|alleged|buttery|present|several|speared|thought|species|involve|remains|bitten|female|attack|caused|killed|little|injury|likely|after|bites|clyde|small|large|dried|cold|post|the|aka|cm|dr|in|kg|lb|to|m|\\/|[a-z0-9\\.]|[\\.\\d\'\"]+|[\\d][\\.]?[\\dmc]{1,2})( |$)"," ",s, perl = TRUE)
  #removed authenticated
  s <- gsub("( to )","",s,perl=TRUE)
  s <- gsub("[^a-z ]+"," ",s,perl=TRUE)
  s <- gsub("( nosed )"," nose ",s, perl = TRUE)
  s <- ifelse(grepl("(leucas)",s, perl = TRUE),"bull shark",s)
  s <- ifelse(grepl("(hoax| not |no shark)",s, perl = TRUE),"no shark",s)
  s <- gsub("whtietip","whitetip",s, perl = TRUE)
  s <- gsub("tawny","tawney",s, perl = TRUE)
  
  s <- gsub("[\\s]+"," ",s, perl = TRUE)
  (str_trim(s))
}

restrictToSharkTerms <- function(s){
  
  if(s=="no shark"){
    return(s)
  }
  
  tokens <- as.list(str_split(s," ")[[1]])
  tokens <- tokens[which(tokens%in%shark_type_terms)]
  s <- str_trim(paste0(tokens,collapse=" "))
  if(nchar(s)>0){
    return(s)
  }
  return(NA)
}

shark_types <- dat2 %>% filter(!is.na(Species)) %>% 
  mutate(shark_type=removeNonTypeWords(Species),
         shark_type=sapply(shark_type,FUN=restrictToSharkTerms),
         shark_type=gsub("^(shark )","",shark_type),
         shark_type=gsub("albimarginatus","silvertip shark",shark_type),
         shark_type=gsub("gangeticus","ganges",shark_type),
         shark_type=ifelse(grepl("(shark)",shark_type),shark_type,paste(shark_type," shark",sep=""))
         )

#write.csv(shark_types,file = "shark_types_2.csv",row.names = FALSE)

all_shark_type_text <- shark_types %>% mutate(shark_type=gsub("shark","",shark_type)) %>% 
  select(shark_type)
all_shark_type_tokens <- str_split(paste(unlist(all_shark_type_text),collapse=" ")," ")
all_shark_type_terms <- as.data.frame(table(term=all_shark_type_tokens))[-1,] %>% filter(!is.na(term)|term!="NA")

all_shark_type_terms %>% arrange(desc(Freq)) %>% kable

wordcloud(all_shark_type_terms$term,all_shark_type_terms$Freq,
            min.freq = 2,c(2.5,1),
            #random.order=FALSE, 
            colors=brewer.pal(6, "Dark2"))

```

####Shark Size

```{r}
feet_pattern <- c("([\\d\\.]+)(\')")
inches_pattern <- c("([\\d\\.]+)(\")")
meters_pattern <- c("([\\d\\.]+)(m)")
cm_pattern <- c("([\\d\\.]+)(cm)")
kg_pattern <- c("([\\d\\.]+)(kg)")
lb_pattern <- c("([\\d\\.]+)(lb)")

size_dat <- dat2 %>%
  mutate(feet=as.numeric(gsub("\'","",stri_extract_last_regex(Species,feet_pattern))),
         inches=as.numeric(gsub("\"","",stri_extract_last_regex(Species,inches_pattern))),
         meters=as.numeric(gsub("m","",stri_extract_last_regex(Species,meters_pattern))),
         cm=as.numeric(gsub("cm","",stri_extract_last_regex(Species,cm_pattern))),
         kg=as.numeric(gsub("kg","",stri_extract_last_regex(Species,kg_pattern))),
         lb=as.numeric(gsub("lb","",stri_extract_last_regex(Species,lb_pattern)))
  )

size_dat <- size_dat %>% 
  mutate(feet=ifelse(!is.na(feet)&!is.na(inches),
                       feet+(inches/12), # combine feet+inches into feet
                     ifelse(is.na(feet)&!is.na(inches),
                            inches/12,feet)),
         meters=ifelse(is.na(meters)&!is.na(feet),feet*0.3048, # 1ft = 0.3048m
                       ifelse(!is.na(cm),cm/100,meters)),
         kg=ifelse(is.na(kg)&!is.na(lb),lb*0.453592,kg) # lb to kg ~ 1 to 0.453592

  ) %>% select(-inches,-cm,-lb)

#write.csv(size_dat,"shark_size.csv",row.names = FALSE,quote=FALSE)

summary(size_dat$meters) #%>% kable(caption='Length Summary (meters)')

summary(size_dat$kg) #%>% kable(caption='Weight Summary (kg)')

```

####Shark Gender

```{r}

gender_dat <- dat2 %>% 
  mutate(Species=gsub("(pregnant)","female",Species),
         gender=ifelse(grepl("(female)",Species),
                       "female",
                       ifelse(grepl("male",Species),
                              "male",
                              NA))
         ) %>% filter(!is.na(gender))
    
   

#write.csv(gender_dat,"shark_gender.csv",row.names = FALSE,quote=FALSE)
gender_dat %>% group_by(gender) %>% summarize(n=n())

```

```{r, eval=FALSE,echo=FALSE,message=FALSE}

# SCRAP THIS CODE CHUNK

#dat2 <- dat %>% mutate(Species=stri_trans_tolower(stringr::str_replace_all(Species,"[\\s,`\\?\\[\\]\\(\\)]+", " ")))

dat2 <- dat %>% 
  mutate(Species=stri_trans_tolower(Species)) # Lowercase all Species text

## Remove all double quotes at the beginning and end of each Species entry
dat2 <- dat2 %>% 
  mutate(Species=sub(pattern="^[\"]","",Species)) %>% #from beginning
  mutate(Species=sub(pattern="([^\\d])[\"]$","\\1",Species)) #ending (unless preceded by a digit)

#dat2 <- dat2 %>% 
#  mutate(Species=gsub(pattern="[,;\\?:-\\[\\]\\(\\)&]"," ",Species)) #doesn't work

dat2 <- dat2 %>% 
  mutate(Species=gsub(pattern="([a-zA-Z])[\\.]","\\1 ",Species)) #removes word-final periods

dat2 <- dat2 %>% mutate(Species=gsub(pattern="\\?+"," ",Species)) #removes question marks

dat2 <- dat2 %>% mutate(Species=gsub(pattern="[,;]"," ",Species)) #removes commas & semicolons

dat2 <- dat2 %>% mutate(Species=gsub(pattern="\\[|\\]"," ",Species)) #removes square brackets

dat2 <- dat2 %>% mutate(Species=gsub(pattern="\\(|\\)"," ",Species)) #remove parentheses

dat2 <- dat2 %>% #normalizes shark length descriptions
  mutate(Species=gsub(pattern="(\\d)(\\s|-)(metre|m|meter|cm|ft|feet|foot)(\\s|$)","\\1\\3 ",Species)) %>%
  mutate(Species=gsub(pattern="(\\d)(?-)(ft|feet|foot)(\\s|$)","\\1' ",Species)) %>% 
  mutate(Species=gsub(pattern="(\\d)(?-)(meter|metre)(\\s|$)","\\1m ",Species)) 

dat2 <- dat2 %>% 
  mutate(Species=gsub(pattern="pound","lb",Species)) #normalizes pound to lb

dat2 <- dat2 %>% #normalizes 'word-tipped' 'wordtipped' 'word tiped' to wordtip
  mutate(Species=gsub(pattern="(\\w)(\\s|-)(tipped)(\\s|$|\")","\\1tip ",Species)) 

## Reduce plural sharks to singular shark

dat2 <- dat2 %>% # normalizes the word 'shark'
  mutate(Species=gsub(pattern="(sharks|shar k|shar ;)","shark",Species)) %>% 
  mutate(Species=gsub(pattern="( shar)\\s+([^k])"," shark \\2",Species)) 
  


dat2 <- dat2 %>% 
  mutate(Species=gsub(pattern="(\\d)(')(shark)","\\1\\2 \\3",Species)) #spaces 5'shark to 5' shark

dat2 <- dat2 %>% 
  mutate(Species=gsub(pattern=" (\")(\\w)"," \\2",Species)) #turns "word into word

dat2 <- dat2 %>% 
  mutate(Species=gsub(pattern="(fragment|wobbegong|day)(s)","\\1",Species))

dat2 <- dat2 %>% 
  mutate(Species=gsub(pattern=" & "," and ",Species))

#%>% filter(!is.na(Species)) %>% select(Species) %>% View() 

#dat2 %>% filter(!is.na(Species)) %>% select(Species) %>% View() 

## Reduce multiple space to a single space
dat2 <- dat2 %>% mutate(Species=gsub(pattern="\\s+"," ",Species))


dat2 %>% filter(!is.na(Species)) %>% select(Species) %>% View() 


#dat2 <- dat2 %>% mutate(Species=stringr::str_replace(Species,"^[\"]",""))

dat2 %>% select(Species) %>% View()
#stri_enc_toutf8

species_counts <- as.data.frame(table(dat2 %>% filter(!is.na(Species)) %>% 
  select(Species)))

species_counts %>% View()


species_tokens <- unlist(str_split(species_counts$Var1," "))
# Remove stopwords
species_tokens <- gsub("(^|\\s)(a|as|by|that|but|their|from|have|an|to|or|be|later|on|was|with|and|the|<U+FFFD>|-|')($|\\s)"," ",species_tokens)

#length(species_tokens)

#species_tokens <- stringr::str_replace(species_tokens,"^[\"]","")
#c("a that but their an to or and the <U+FFFD> - & '")
species_token_counts <- as.data.frame(table(species_tokens))
#length(species_token_counts$species_tokens[[2]]) # %>% filter(length(species_tokens)>1) %>% View()

#filter stopwords...
#stop_words_species <- unlist(str_split("a that their it but an to or and the <U+FFFD> - ' &"," "))
#stop_words_species <- str_split(c("a that but their an to or and the <U+FFFD> - & '")," ")
#species_token_counts <- species_token_counts %>% filter(!species_tokens%in%stop_words_species)
#densityplot(x=species_token_counts$Freq,data=species_token_counts) + scale_y_log10()



wordcloud(species_token_counts$species_tokens,
          species_token_counts$Freq,min.freq = 9,#c(4,1),
          random.order=FALSE, 
          colors=brewer.pal(8, "Dark2")) #ordered.colors=TRUE,colors = TRUE)



```


##Wrangle the Activity column

```{r, warning=FALSE}
#Wrangle the activity column..

#names(dat)[names(dat)=="Case Number"] <- "CaseNumber"

activity_dat <- dat %>% select(`Case Number`,Activity) %>% filter(!is.na(Activity))

activity_dat <- activity_dat %>% 
  mutate(Activity=iconv(Activity,from="CP1252",to="UTF-8"),
         Activity=stri_trans_tolower(Activity)) # Lowercase all Activity text

# Remove annoying unicode characters
activity_dat <- activity_dat %>%
  mutate(Activity=gsub("(\u393C|\u3E33|\u613C|\u3E30|\u3E34|\u623C|\u3E64|\uFFFD)","",Activity))

activity_dat <- activity_dat %>% mutate(Activity=str_trim(gsub("[\\s\"'`\\?\\(\\),;:-]+"," ",Activity,perl=TRUE))) #%>% View()
#activity_dat$Activity <- stringr::str_replace_all(activity$Activity,"[\\s\"'`\\?\\(\\)]+", " ")

activity_dat %>% group_by(Activity) %>% summarize(freq=n()) %>% arrange(desc(freq)) %>% head(20)

activity_dat %>% filter(grepl("ing( |$)",Activity)) %>% 
  group_by(Activity) %>% summarize(freq=n()) %>% arrange(desc(freq)) %>% head(20)

# How deep?
activity_dat %>% filter(grepl("deep",Activity)) %>% 
  mutate(depth=str_extract(Activity,"([^\\s]+[\\s]){2}(deep|depth)")) %>% 
  group_by(depth) %>% summarize(freq=n()) %>% arrange(desc(freq)) %>% head(20)


activity_tokens <- unlist(str_split(activity_dat$Activity," "))
num_tok <- length(activity_tokens)
num_uniq <- length(unique(activity_tokens))
act_word_counts <- as.data.frame(table(term=activity_tokens))

act_stop_words <- c("another", "been", "a", "as", "an", "by", "of", "or", "on", "that", "his", "but", "it", "its", "their", "than", "from", "have", "out", "were", "be", "is", "was", "later", "on", "with", "and", "the")

act_word_counts %>% filter(!term%in%act_stop_words) %>% head(20)


wordcloud(act_word_counts$term,act_word_counts$Freq,
          min.freq = 8,
          random.order=FALSE, 
          colors=brewer.pal(8, "Dark2"))


```


